Fire danger: the skill provided by the ECMWF Integrated Forecasting System

### 2020-05-20 scripts/article.Rmd
Removed files connected with article.rmd

In addition, from ERA5 we derive a climatological benchmark simulation (called CLIM hereafter), considered as a "mean year" for the FWI using the period 1980-2018. At pixel level, every day the median FWI is calculated over the period spanning 4 days before/after a given date. While CLIM has no expected predictive skill for a specific year, it retains information on the yearly climatological variation of FWI. CLIM is what could be used in the absence of either a monitoring (based on observations) or a forecast system (based on weather forecast). It should therefore score better or equal of the forecast on time ranges beyond the limits of predictability. CLIM is used in this study as a benchmark to rank the expected improvements provided by a forecasting system

@@

# Load relevant packages
library("caliver")
library("raster")
library("ggplot2")
library("verification")
library("rowr")
library("dplyr")
library("rgeos")
library("maptools")
library("sp")
library("RColorBrewer")
library("stringr")
library("colorspace")
library("easyVerification")
library("spatialEco")

forecasts_folder <- "/hugetmp/forecasts/fwi"

@@

# Get all GFED4 regions
BasisRegions <- readRDS("data/BasisRegions_simplified.rds")
```

@@

```{r prep, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
# The following code uses forecast data which is restricted.
# A summary table is computed and saved for users at the end of this code chunk.
#
# Get data from previous publication
# https://github.com/cvitolo/GEFF-ERA5/blob/master/data/df_geff_erai_era5.rds
# save it in the "../data" folder
# df <- readRDS(file = "data/df_geff_erai_era5.rds")
# # insert 1-year fire season for tropics (lat = [-30, +30])
# df$season[df$lat < 30 & df$lat > -30] <- "Dry"
# # Remove wet season (we are only interested in the dry season)
# df <- df[df$season == "Dry", ]
# # How many stations remain?
# stations <- unique(df[, c("id", "lat", "long")])
# 
# df$region <- sapply(strsplit(df$tzid, "/"), `[`, 1)
# df$date <- as.Date(paste0(df$yr, "-", df$mon, "-", df$day))
# df <- df[, c("id", "lat", "long", "region", "date", "OBS", "ERA5")]
# 
# df_used <- data.frame(matrix(NA, nrow = 0, ncol = 37))
# names(df_used) <- c("id", "lat", "long", "region", "date", "OBS", "ERA5",
#                     paste0(c("CLIM_d", "HRES_d", "CRPS_d"), rep(1:10, each =3)))
# 
# for (i in seq_along(dates2017)){
# 
#   print(dates2017[i])
# 
#   # What stations have data on this date?
#   dfx <- df %>% filter(date == dates2017[i]) %>% filter(complete.cases(.))
#   # Define spatial point
#   spdf <- dfx[, c("id", "lat", "long")]
#   # Promote data.frame to spatial
#   coordinates(spdf) = ~long + lat
#   crs(spdf) <- new("CRS",
#                    projargs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
#   # Get data from ENS forecasts
#   nstations <- dim(spdf@data)[1]
# 
#   # Get data from HRES forecasts, e.g. ECMWF_FWI_20170101_1200_hr_fwi.nc
#   for (leadtime in 1:10){
# 
#     lead_day <- dates2017[i] + leadtime - 1
# 
#     # Get ERA5 data
#     message(paste("Handling ERA5 for", lead_day))
#     dayx <- which(format(datesERA5, "%m-%d") %in% format(lead_day, "%m-%d")
#                   & lubridate::year(datesERA5) != "2017")
#     system(paste0("cdo seltimestep,", paste0(dayx, collapse = ","),
#                   " /scratch/rd/nen/perClaudia/era5/fwi_1980_2019.nc ",
#                   "/scratch/rd/nen/perClaudia/era5/fwi_1980_2019_dayx.nc"),
#            ignore.stderr = TRUE)
#     ERA5 <- brick("/scratch/rd/nen/perClaudia/era5/fwi_1980_2019_dayx.nc")
#     # ERA5 was stored with longitudes already rotated to [-180, +180],
#     # HRES and ENS will need to be rotated
#     df_era5 <- raster::extract(x = ERA5, y = spdf)
# 
#     # Get HRES
#     message(paste("Handling HRES for", dates2017[i], "day", leadtime))
#     HRES <- raster::brick(file.path(forecasts_folder,
#                                     paste0("ECMWF_FWI_",
#                                            gsub("-", "",
#                                                 as.character(dates2017[i])),
#                                            "_1200_hr_fwi.nc")))[[leadtime]]
#     df_hres <- raster::extract(x = rotate(HRES), y = spdf)
# 
#     # Get ENS
#     df_ens <- matrix(NA, nrow = nstations, ncol = 51)
#     message(paste("Handling ENS for", dates2017[i], "day", leadtime))
#     for (j in 1:51){
#       # Extract the modelled FWI from ENS
#       ens_member <- sprintf("%02d", j - 1)
#       ENS <- raster::brick(file.path(forecasts_folder,
#                                      paste0("ECMWF_FWI_",
#                                             gsub("-", "",
#                                                  as.character(dates2017[i])),
#                                             "_1200_", ens_member,
#                                             "_fwi.nc")))[[leadtime]]
#       df_ens[, j] <- raster::extract(x = rotate(ENS), y = spdf)
#     }
# 
#     # Remove rows with NAs and update spatial points
#     if (leadtime == 1){
#       rows2remove <- which(rowSums(is.na(df_era5)) > 0)
#       if (length(rows2remove) > 0){
#         df_ens <- df_ens[-rows2remove, ]
#         df_hres <- df_hres[-rows2remove]
#         df_era5 <- df_era5[-rows2remove, ]
#         # Resample ERA5 to get 39 random ensemble members
#         df_era5 <- df_era5[, sample(x = 1:dim(df_era5)[2], size = 39,
#                                     replace = TRUE)]
# 
#         dfx <- dfx[-rows2remove, ]
#         # Re-define spatial point
#         spdf <- dfx[, c("id", "lat", "long")]
#         # Promote data.frame to spatial
#         coordinates(spdf) = ~long + lat
#         crs(spdf) <- new("CRS",
#                          projargs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
#         # Get data from ENS forecasts
#         nstations <- dim(spdf@data)[1]
#       }
#     }
# 
#     # Populate CRPS
#     df_crps <- veriApply(verifun = "EnsCrps", fcst = df_ens, obs = df_era5)
#     df_crpss <- apply(X = df_crps, MARGIN = 1, FUN = mean)
#     df_era5_mean <- apply(X = df_era5, MARGIN = 1, FUN = mean)
# 
#     dfx <- cbind(dfx, df_era5_mean, df_hres, df_crpss)
#   }
# 
#   names(dfx) <- names(df_used)
# 
#   # Append to table
#   df_used <- dplyr::bind_rows(df_used, dfx)
#   saveRDS(df_used, "data/df_geff_era5_hres_crps.rds")
# }

df_used <- readRDS("../data/df_geff_era5_hres_crps.rds")

# Promote data.frame to spatial
coordinates(df_used) = ~long + lat

# Countries of interest
pt <- raster::getData(name = "GADM", country = "Portugal", level = 1)[-c(2,13),]
chile <- raster::getData(name = "GADM", country = "Chile", level = 0)
cali <- raster::getData(name = "GADM", country = "USA", level = 1)[5, ]

# Synop stations of interest
crs(df_used) <- crs(BasisRegions)
synops <- spatialEco::point.in.poly(df_used, BasisRegions)
# saveRDS(synops, "../data/synops.rds")
```

@@

```{r Figure3, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure4}PUT here crps skills scores."}
f3 <- data.frame(synops) %>%
  select(Region, CRPS_d1, CRPS_d2, CRPS_d3, CRPS_d4, CRPS_d5, CRPS_d6,
         CRPS_d7, CRPS_d8, CRPS_d9, CRPS_d10) %>%
  reshape2::melt(id.vars = "Region") %>%
  filter(complete.cases(.)) %>%
  rename(CRPS = value) %>%
  group_by(Region, variable) %>%
  summarise(CRPS = mean(CRPS))

@@

@@

```{r prepFig4, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
# Get ERA5
ERA5 <- brick("/scratch/rd/nen/perClaudia/era5/fwi_1980_2019.nc")
# Get FRP
FRP <- raster::brick("/scratch/rd/nen/perClaudia/CAMS/CAMS_2017-01-01_2017-12-31_frpfire.nc")
myfilelist <- list.files(path = forecasts_folder,
                        pattern = "*hr",
                        full.names = TRUE)

library(doSNOW)
ncores <- parallel::detectCores() - 1
cl <- makeCluster(ncores)
registerDoSNOW(cl)
iterations <- 365
pb <- txtProgressBar(max = iterations, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
result <- foreach(i = 1:iterations, .combine = rbind, 
                  .options.snow = opts) %dopar%
{
    # Get FRP for 2017, keep only major fires (values above 0.5)
  frpx <- raster::rotate(FRP[[i]])
  p <- raster::rasterToPoints(frpx, fun = function(x){x > 0.5})
  spdf <- data.frame(p)
  spdf$date <- dates2017[i]
  names(spdf) <- c("long", "lat", "frp", "date")
  sp::coordinates(spdf) <- ~long + lat

  # Get HRES for 2017, where FRP is above 0.5
  hres <- raster::rotate(raster::brick(myfilelist[[i]]))
  x <- raster::extract(x = hres, y = spdf)
  x <- data.frame(x)
  names(x) <- paste0("Day", 1:10)

  # Get ERA5 to calculate the very high warning levels
  very_high <- caliver::daily_clima(r = ERA5, dates = dates2017[i], probs = 0.95)
  thr95 <- raster::extract(x = very_high, y = spdf)

  # Append results to table
  df <- cbind(spdf@coords, spdf@data, x, thr95)
  df <- df[complete.cases(df), ]
  return(df)
}
close(pb)
stopCluster(cl)
# saveRDS(result, "data/PODdataframe.rds")

# FRP locations of interest
sp::coordinates(result) <- ~long + lat
crs(result) <- crs(BasisRegions)
fires <- spatialEco::point.in.poly(result, BasisRegions)
# saveRDS(fires, "data/fires.rds")
```

```
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{POD2.png}\\[0.5cm]
\includegraphics[width=0.5\textwidth]{POD6.png}\\[0.5cm]
\includegraphics[width=0.5\textwidth]{POD10.png}\\[0.5cm]
\includegraphics[width=0.5\textwidth]{PODclima.png}\\[0.5cm]
\caption{Global area averaged Probability of Detection (POD) for  day 2, day 6 and day 10 forecasts and the CLIM run. Pixels where FRP $ \ge $ 0.5 Wm$^{-2}$  are categorized as "yes" events and compared to FWI prediction above the very-high warning level. The global statistic is constructed using all FRP observations detected in 2017 and averaged over the specified regions.} \label{fig:pod}
\end{figure}
```

@@

```
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Chile_boxy_map}
        \caption{Chile}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Portugal_boxy_map}
        \caption{Portugal}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{California_boxy_map}
        \caption{California}
    \end{subfigure}
    \caption{Comparison of Fire Radiative Power (gray dashed line with axis on the right hand side) with FWI forecasted using the deterministic high resolution model for California, Portugal and  Chile. FWI is color coded based on the percentage of pixels exceeding the high danger level calculated at the country/state level. Each of the panel refers to a specific fire event described in the text and the statistics have been calculated over the red boxes.} \label{fig:plotthatrulesthemall}
\end{figure}
```

One of the advantages of developing a probabilistic ensemble prediction system (ENS) for fire danger is the possibility to assess the confidence in the model forecast, at least from the point of view of the uncertainties in weather forcings. Admittedly, the use of this type of information is not straightforward as many studies have highlighted \citep{pappenberger:13,palmer:00}. This difficulty steams from the fact that in many response systems an activation is conditional to the excedance of  deterministic threshold. Instead, in a probabilistic approach what is provided is the {\it probability} of exceedance a certain values and this  is often perceived as a degradation of the information values \citep{richardson:00}. However by using the probabilistic component of the system, it is possible to understand how consistent the fire danger forecast is with respect to slightly different forecast scenarios. To highlight the information that can be added with the use of the ensemble prediction system, Figure \ref{fig:fire_ensfwi} shows the progression of 10-day forecasts for the major fire events (corresponding to the peak of observed FRP in figure \ref{fig:plotthatrulesthemall}) in the three case studies. In this case we concentrate on a specific location (model grid box) whose coordinates are specified in table \ref{tab:fires}. In this plot, there is very high consistency between HRES and ENS, with the former falling generally close to the ensemble median. In addition, the bulk of ensemble members falls in the yellow region which identifies very high danger. It is this consistency and persistence of very high danger over time that could have helped users gain confidence in the forecast, had that been already available. While these events were all considered catastrophic in terms of impacts, it is noticeable that, at all forecast ranges, fire danger would have been classified as very-high but not extreme by the exclusive use of the high resolution forecast. However in  all three cases by looking at the full distribution of FWI values provided by the ensemble, it is clear that some of the forecasts were hinting at the possibility for extreme conditions.

Thus is there an enhanced accuracy in the forecast produced with the use of the ensemble forecasting system? Figure \ref{fig:fire_enspod} shows the probability of detection calculated over the three fire regions (see red box in Figure \ref{fig:plotthatrulesthemall} for the geographical extents of these areas) using all the ensemble members. The high resolution forecast (red dots) is not in all cases the most skillful prediction despite the increased resolution. It is interesting to notice how in the California case, for example, many of the ensemble members perform consistently better than the high resolution run up to 6 days ahead.

>>>>

In addition, from ERA5 we derive a climatological benchmark simulation (called CLIM hereafter), considered as a "mean year" for the FWI using the period 1980-2018. At pixel level, every day the median FWI is calculated over the period spanning 4 days before/after a given date. While CLIM has no expected predictive skill for a specific year, it retains information on the yearly climatological variation of FWI. CLIM is what could be used in the absence of either a monitoring (based on observations) or a forecast system (based on weather forecast). It should therefore score better or equal of the forecast on time ranges beyond the limits of predictability. CLIM is used in this study as a benchmark to rank the expected improvements provided by a forecasting system.

@@

# Load relevant packages
library("caliver")
library("raster")
library("ggplot2")
library("rowr")
library("dplyr")
library("rgeos")
library("maptools")
library("sp")
library("sf")
library("RColorBrewer")
library("stringr")
library("colorspace")
library("easyVerification")
library("spatialEco")
library("doSNOW")
library("parallel")
library("gplots")

setwd("data")
forecasts_folder <- "/hugetmp/forecasts/hres/2017"

@@

# Get all GFED4 regions
BasisRegions <- readRDS("BasisRegions_simplified.rds")

# Countries of interest
pt <- raster::getData(name = "GADM", country = "Portugal", level = 1)[-c(2,13),]
chile <- raster::getData(name = "GADM", country = "Chile", level = 0)
cali <- raster::getData(name = "GADM", country = "USA", level = 1)[5, ]
```

@@

```{r prep, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE, eval = FALSE}
# The following code uses forecast data which is restricted.
# A summary table is computed and saved for users at the end of this code chunk.
#
# Get data from previous publication
# https://github.com/cvitolo/GEFF-ERA5/blob/master/data/df_geff_erai_era5.rds
# save it in the "../data" folder
df <- readRDS(file = "df_geff_erai_era5.rds")
# insert 1-year fire season for tropics (lat = [-30, +30])
df$season[df$lat < 30 & df$lat > -30] <- "Dry"
# Remove wet season (we are only interested in the dry season)
df <- df[df$season == "Dry", ]
# How many stations remain?
stations <- unique(df[, c("id", "lat", "long")])

df$region <- sapply(strsplit(df$tzid, "/"), `[`, 1)
df$date <- as.Date(paste0(df$yr, "-", df$mon, "-", df$day))
df <- df[, c("id", "lat", "long", "region", "date", "OBS", "ERA5")]

days_2017 <- which(datesERA5 %in% c(dates2017, tail(dates2017)[6] + 1:9))
# system(paste0("cdo seltimestep,", paste0(days_2017, collapse = ","),
#                   " /scratch/rd/nen/perClaudia/era5/fwi_1980_2019.nc ",
#                   "/scratch/rd/nen/perClaudia/era5/fwi_2017_dayx.nc"),
#            ignore.stderr = TRUE)
ERA5_2017 <- raster::brick("/scratch/rd/nen/perClaudia/era5/fwi_2017_dayx.nc")

for (i in seq_along(dates2017)) {

  issue_date <- dates2017[i]

  # What stations have data on this date?
  dfx <- dplyr::filter(.data = df, date == issue_date)
  dfx <- dfx[complete.cases(dfx), ]

  # Define spatial point
  spdf <- sf::st_as_sf(x = dfx, coords = c("long", "lat"), crs = 4326)

  # Get HRES
  message(paste("Handling HRES for", issue_date))
  HRES <- raster::brick(file.path(forecasts_folder,
                                  paste0("ECMWF_FWI_",
                                         gsub("-", "",
                                              as.character(issue_date)),
                                         "_1200_hr_fwi.nc")))
  # HRES needs to be rotated
  df_hres <- raster::extract(x = raster::rotate(HRES), y = spdf)
  dfx <- cbind(dfx, df_hres)
  names(dfx)[8:17] <- paste0("HRES_d", 1:10)

  # Get ENS
  arr_ens <- array(NA, dim = c(dim(spdf)[1], 10, 51))
  message(paste("Handling ENS for", issue_date))
  for (j in 1:51){
    # Extract the modelled FWI from ENS
    ens_member <- sprintf("%02d", j - 1)
    ENS <- raster::brick(file.path(forecasts_folder,
                                   paste0("ECMWF_FWI_",
                                          gsub("-", "",
                                               as.character(issue_date)),
                                          "_1200_", ens_member,
                                          "_fwi.nc")))[[1:10]]
    arr_ens[, , j] <- raster::extract(x = raster::rotate(ENS), y = spdf)
  }

  # Get ERA5 data for 2017 - to be used as observation
  message(paste("Handling ERA5 for", issue_date, "plus 9 days"))
  ERA5_2017_day <- ERA5_2017[[which(dates2017 %in% issue_date) + 0:9]]
  day_in_2017 <- raster::extract(x = ERA5_2017_day, y = spdf)

  for (leadtime in 1:10){

    lead_day <- issue_date + leadtime - 1

    # Extract ERA5 from all the years (same month and day), excluding 2017
    idx_in_datesERA5 <- which(format(datesERA5, "%m-%d") %in%
                              format(lead_day, "%m-%d") &
                              lubridate::year(datesERA5) != "2017")
    system(paste0("cdo seltimestep,", paste0(idx_in_datesERA5, collapse = ","),
                  " /scratch/rd/nen/perClaudia/era5/fwi_1980_2019.nc ",
                  "/scratch/rd/nen/perClaudia/era5/fwi_1980_2019_lead_day.nc"),
           ignore.stderr = TRUE)
    ERA5 <- raster::brick("/scratch/rd/nen/perClaudia/era5/fwi_1980_2019_lead_day.nc")

    # ERA5 was stored with longitudes already rotated to [-180, +180],
    # HRES and ENS will need to be rotated
    df_era5_39 <- raster::extract(x = ERA5, y = spdf)
    # Sample ERA5 to get 51 random ensemble members
    df_era5_51 <- df_era5_39[, sample(x = 1:dim(df_era5_39)[2],
                                      size = 51, replace = TRUE)]

    # Populate CRPS
    crps_fc <- easyVerification::veriApply(verifun = 'EnsCrps',
                         fcst = arr_ens[, leadtime,],
                         obs = day_in_2017[, leadtime])
    crps_clim <- easyVerification::veriApply(verifun = 'EnsCrps',
                           fcst = df_era5_51,
                           obs = day_in_2017[, leadtime])
    crpss <- 1 - crps_fc/crps_clim

    dfx <- cbind(dfx, crps_fc, crps_clim, crpss)
  }
  names(dfx)[18:47] <- paste0(c("CRPS_fc_d", "CRPS_clim_d", "CRPSS_d"),
                              rep(1:10, each = 3))

  if (i == 1) {
    result <- dfx
  }else{
    result <- rbind(result, dfx)
  }

  saveRDS(result, "df_geff_era5_hres_crps.rds")

}
```

```{r prep2, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
result <- readRDS("df_geff_era5_hres_crps.rds")

# Remove rows with NAs and update spatial points
rows2remove <- which(rowSums(is.na(result)) > 0)
if (length(rows2remove) > 0) result <- result[-rows2remove, ]

# Define spatial point
spdf <- sf::st_as_sf(x = result, coords = c("long", "lat"), crs = 4326)

synops <- spatialEco::point.in.poly(spdf, BasisRegions)
```

@@

```{r Figure3, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure4}PUT here crps skills scores."}
f3 <- data.frame(synops) %>%
  select(Region, CRPS_fc_d1, CRPS_fc_d2, CRPS_fc_d3, CRPS_fc_d4, CRPS_fc_d5, CRPS_fc_d6,
         CRPS_fc_d7, CRPS_fc_d8, CRPS_fc_d9, CRPS_fc_d10) %>%
  reshape2::melt(id.vars = "Region") %>%
  filter(complete.cases(.)) %>%
  rename(CRPS = value) %>%
  group_by(Region, variable) %>%
  summarise(CRPS = mean(CRPS))

@@

```{r Figure4, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure4}PUT here crps skills scores."}
f4 <- data.frame(synops) %>%
  filter(complete.cases(.)) %>%
  select(paste0(c("CRPS_fc_d", "CRPS_clim_d"), rep(1:10, each = 2))) %>%
  reshape2::melt(id.vars = c("CRPS_clim_d1", "CRPS_clim_d2", "CRPS_clim_d3",
                             "CRPS_clim_d4", "CRPS_clim_d5", "CRPS_clim_d6",
                             "CRPS_clim_d7", "CRPS_clim_d8", "CRPS_clim_d9",
                             "CRPS_clim_d10")) %>%
  rename(CRPS_fc = value) %>%
  rename(fc_name = variable) %>%
  reshape2::melt(id.vars = c("fc_name", "CRPS_fc")) %>%
  rename(CRPS_clim = value) %>%
  rename(clim_name = variable) %>%
  group_by(fc_name, clim_name) %>%
  summarise(CRPS_fc = mean(CRPS_fc),
            CRPS_clim = mean(CRPS_clim))

f4$fc_name <- as.numeric(f4$fc_name)
f4$clim_name <- as.numeric(f4$clim_name)

f4plot <- ggplot(f4) +
  geom_line(aes(x = fc_name, y = CRPS_fc, col = "Forecast"), size = 2) +
  geom_line(aes(x = clim_name, y = CRPS_clim, col = "Climatology"), size = 2) +
  xlab("Lead time") + ylab("CRPS") +
  scale_color_discrete(name = "") +
  theme_bw() +
  scale_x_discrete(limits = 1:10) +
  theme(text = element_text(size = 15))

ggsave(filename = "../images/Figure4.png", plot = f4plot)
```

```{r Figure5, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure4}PUT here crps skills scores."}
f5 <- data.frame(synops) %>%
  filter(complete.cases(.)) %>%
  select(Region, paste0(c("CRPS_fc_d", "CRPS_clim_d"), rep(1:10, each = 2))) %>%
  reshape2::melt(id.vars = c("Region",
                             "CRPS_clim_d1", "CRPS_clim_d2", "CRPS_clim_d3",
                             "CRPS_clim_d4", "CRPS_clim_d5", "CRPS_clim_d6",
                             "CRPS_clim_d7", "CRPS_clim_d8", "CRPS_clim_d9",
                             "CRPS_clim_d10")) %>%
  rename(CRPS_fc = value) %>%
  rename(fc_name = variable) %>%
  reshape2::melt(id.vars = c("Region", "fc_name", "CRPS_fc")) %>%
  rename(CRPS_clim = value) %>%
  rename(clim_name = variable) %>%
  group_by(Region, fc_name, clim_name) %>%
  summarise(CRPS_fc = mean(CRPS_fc),
            CRPS_clim = mean(CRPS_clim))

f5$fc_name <- as.numeric(f5$fc_name)
f5$clim_name <- as.numeric(f5$clim_name)
f5$Region <- factor(f5$Region)

ggplot(f5) +
  geom_line(aes(x = fc_name, y = CRPS_fc, col = "Forecast")) +
  geom_line(aes(x = clim_name, y = CRPS_clim, col = "Climatology")) +
  facet_wrap( ~ Region, ncol = 2, scales = "free_y") +
  xlab("Lead time") + ylab("CRPS") +
  scale_color_discrete(name = "") +
  theme_bw() +
  scale_x_discrete(limits = 1:10)
```

@@

```{r prepFig6_7, eval = FALSE, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
# Get ERA5
ERA5 <- raster::brick("/scratch/rd/nen/perClaudia/era5/fwi_1980_2019.nc")
# Get FRP
FRP <- raster::brick("/scratch/rd/nen/perClaudia/CAMS/CAMS_2017-01-01_2017-12-31_frpfire.nc")
myfilelist <- list.files(path = forecasts_folder,
                        pattern = "*hr",
                        full.names = TRUE)

ncores <- parallel::detectCores() - 1
cl <- makeCluster(ncores)
registerDoSNOW(cl)
iterations <- 365
pb <- txtProgressBar(max = iterations, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
result <- foreach(i = 10:iterations, .combine = rbind, 
                  .options.snow = opts) %dopar%
{
  # Get FRP for 2017, keep only major fires (values above 0.5)
  frpx <- raster::rotate(FRP[[i]])
  p <- raster::rasterToPoints(frpx, fun = function(x){x > 0.5})
  spdf <- data.frame(p)
  spdf$date <- dates2017[i]
  names(spdf) <- c("long", "lat", "frp", "date")
  sp::coordinates(spdf) <- ~long + lat
  n_points <- dim(spdf@data)[1]

  # Get HRES for 2017, where FRP is above 0.5
  lead_day <- 0
  hres_df <- data.frame(matrix(NA, ncol = 10, nrow = n_points))
  for (j in i:(i - 9)){
    lead_day <- lead_day + 1
    hres <- raster::rotate(raster::brick(myfilelist[[j]])[[lead_day]])
    x <- raster::extract(x = hres, y = spdf)
    hres_df[, lead_day] <- x
  }
  names(hres_df) <- paste0("hres_day", 1:10)

  # Get ERA5 to calculate various warning levels
  # Low = 0.75, Moderate = 0.85, High = 0.90, Very high = 0.95, Extreme = 0.98
  median <- caliver::daily_clima(r = ERA5, dates = dates2017[i], probs = 0.50)
  moderate <- caliver::daily_clima(r = ERA5, dates = dates2017[i], probs = 0.85)
  high <- caliver::daily_clima(r = ERA5, dates = dates2017[i], probs = 0.90)
  very_high <- caliver::daily_clima(r = ERA5, dates = dates2017[i], probs = 0.95)
  CLIM <- raster::extract(x = median, y = spdf)
  thr85 <- raster::extract(x = high, y = spdf)
  thr90 <- raster::extract(x = high, y = spdf)
  thr95 <- raster::extract(x = very_high, y = spdf)

  # Append results to table
  df <- cbind(spdf@coords, spdf@data, CLIM, thr85, thr90, thr95, hres_df)
  df <- df[complete.cases(df), ]
  return(df)
}
close(pb)
stopCluster(cl)
# saveRDS(result, "PODdataframe.rds")

# Load, if pre-calculate
# result <- readRDS("PODdataframe.rds")
# BasisRegions <- readRDS("BasisRegions_simplified.rds")

# FRP locations of interest
sp::coordinates(result) <- ~long + lat
crs(result) <- crs(BasisRegions)

# Add ID, Region and Zone to the dataset vias spatial intersection
fires <- spatialEco::point.in.poly(result, BasisRegions)
fires <- as.data.frame(fires)
fires <- fires[complete.cases(fires), ]

# Check coverage
# plot(BasisRegions)
# plot(result, col = "red", add = TRUE)

# If all looks good, we can remove previous datasets
# rm(result, BasisRegions)

# saveRDS(fires, "fires.rds")
```

```{r prepFig6, eval = FALSE, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
# fires <- readRDS("fires.rds")

# Check if CLIM and HRES are above threshold (90th percentile)
my_thr <- fires$thr90
fires$CLIM <- fires$CLIM >= my_thr
for (colx in which(names(fires) %in% paste0("hres_day", 1:10))){
  x <- fires[, colx] > my_thr
  fires <- cbind(fires, x)
}
names(fires)[which(names(fires) == "x")] <- paste0("HDay", 1:10)
rm(x, colx)

# Group by Region and calculate POD
fires_binary <- fires %>%
  mutate(FRP_binary = TRUE) %>%
  select(Region, coords.x1, coords.x2, FRP_binary, CLIM,
         HDay1, HDay2, HDay3, HDay4, HDay5,
         HDay6, HDay7, HDay8, HDay9, HDay10) %>%
  reshape2::melt(id.vars = c("Region", "coords.x1", "coords.x2", "FRP_binary")) %>%
  group_by(Region, coords.x1, coords.x2, variable) %>%
  summarise(POD = verify(obs = FRP_binary, pred = value,
                         frcst.type = "binary", obs.type = "binary")$POD) %>%
  group_by(Region, variable) %>%
  summarise(PODmean = mean(POD))
# saveRDS(fires_binary, "fires_binary_90.rds")
```

```{r Figure6, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure5}Global area averaged Probability of Detection (POD) for  day 2, day 6 and day 10 forecasts and the CLIM run. Pixels where FRP $ \ge $ 0.5 Wm$^{-2}$  are categorized as "yes" events and compared to FWI prediction above the very-high warning level. The global statistic is constructed using all FRP observations detected in 2017 and averaged over the specified regions."}
fires_binary <- readRDS("fires_binary_90.rds")
fires_binary <- fires_binary[-which(fires_binary$variable == "CLIM"), ]

ggplot(fires_binary, aes(variable, Region, fill = PODmean)) + 
  geom_tile() +
  geom_text(aes(label = round(PODmean, 2))) +
  scale_fill_gradient(name = "POD", low = "white", high = "brown") +
  theme_bw() + xlab("") + ylab("")
```

```{r Figure7, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure4}GFED4 regional classification and the locations of events reported as Fire Radiative Power (FRP) by GFAS."}
ggplot(data = map_data("world"), aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill=NA, colour = "grey65") +
  coord_equal() +  theme_bw() + xlab("Longitude") + ylab("Latitude") +
  geom_polygon(data = regionXPoly,
               aes(x = long, y = lat, group = group, fill = factor(Region)),
               alpha = 0.5) +
  scale_fill_manual(name="GFED4 regions",
                    values = c("darkgrey",
                               RColorBrewer::brewer.pal(12,"Paired"),
                               "lightgrey"),
                    labels = BasisRegions$Region) +
  geom_point(data = data.frame(result), aes(x = long, y = lat),
             shape = 1, color = "gray20", size = 0.1)
```

@@

```{r prepFig6_maps, eval = FALSE, echo = FALSE, message = FALSE, results = 'hide', cache = TRUE}
# Chile
dates_chile <- seq.Date(from = as.Date("2017-01-01"),
                        to = as.Date("2017-01-31"),
                        by = "day")
bbox_chile <- as(raster::extent(-74, -70, -37, -33), "SpatialPolygons")
raster::plot(chile, col = "lightgray")
raster::plot(bbox_chile, lwd = 2, border = "red", add = TRUE)
# Crop manually to mainland!

# Portugal
dates_pt <- seq.Date(from = as.Date("2017-06-01"),
                     to = as.Date("2017-06-30"),
                     by = "day")
bbox_pt <- as(raster::extent(-8.5, -7.6, 39.6, 40.3), "SpatialPolygons")
pt <- gUnaryUnion(pt, id = pt@data$NAME_0)
raster::plot(pt, col = "lightgray")
raster::plot(bbox_pt, lwd = 2, border = "red", add = TRUE)
# Crop manually to mainland!

# California
dates_cali <- seq.Date(from = as.Date("2017-09-21"),
                       to = as.Date("2017-10-20"),
                       by = "day")
bbox_cali <- as(raster::extent(-123.4, -117.6, 33.7, 42), "SpatialPolygons")
raster::plot(cali, col = "lightgray")
raster::plot(bbox_cali, lwd = 2, border = "red", add = TRUE)
# Crop manually to mainland!

obs_file <- "/scratch/rd/nen/perClaudia/CAMS/CAMS_2017-01-01_2017-12-31_frpfire.nc"
clima_file <- "/scratch/rd/nen/perClaudia/era5/fwi_era5_1980_2016_90th_daily_clima.nc"

# Chile
df_chile <- caliver:::make_forecast_summary(input_dir = "/hugetmp/forecasts/fwi",
                                            p = bbox_chile,
                                            event_dates = dates_chile,
                                            obs = obs_file,
                                            clima = clima_file)
saveRDS(df_chile, "df_chile.rds")
# Portugal
df_pt <- caliver:::make_forecast_summary(input_dir = "/hugetmp/forecasts/fwi",
                                         p = bbox_pt,
                                         event_dates = dates_pt,
                                         obs = obs_file,
                                         clima = clima_file)
saveRDS(df_pt, "df_pt.rds")
# California
df_cali <- caliver:::make_forecast_summary(input_dir = "/hugetmp/forecasts/fwi",
                                           p = bbox_cali,
                                           event_dates = dates_cali,
                                           obs = obs_file,
                                           clima = clima_file)
saveRDS(df_cali, "df_cali.rds")
```

```{r Figure6, echo = FALSE, fig.width = 13, fig.height = 6, out.width = "100%", fig.cap = "\\label{fig:Figure6a}Chile."}
# Chile
plot_chile <- caliver:::plot_forecast_summary(df_chile)
# Portugal
# df_pt <- readRDS("df_pt.rds")
plot_pt <- caliver:::plot_forecast_summary(df_pt)
# California
# df_cali <- readRDS("df_cali.rds")
plot_cali <- caliver:::plot_forecast_summary(df_cali)

#\caption{Comparison of Fire Radiative Power (gray dashed line with axis on the right hand side) with FWI forecasted using the deterministic high resolution model for California, Portugal and  Chile. FWI is color coded based on the percentage of pixels exceeding the high danger level calculated at the country/state level. Each of the panel refers to a specific fire event described in the text and the statistics have been calculated over the red boxes.} \label{fig:plotthatrulesthemall}
```
